Different rules for binocular combination of luminance flicker in cortical and subcortical pathways

### 2023-06-26 manuscript.Rmd
Simplified presentation of model figure

text(9,0.58,'(i)',cex=1.25,adj=0.5)

@@

legend(21,0.04,c('Monocular','Binocular','Dichoptic','Dichoptic cross'),lwd=3,col=colpal[condlist],box.lwd=2)

text(17,0.037,'(e)',adj=0.5,cex=2.5)

@@

# legend(21,0.8,c('Monocular','Binocular','Dichoptic','Dichoptic cross'),lwd=3,col=colpal[condlist],box.lwd=2)

text(17,0.95,'(f)',adj=0.5,cex=2.5)

@@

# legend(21,0.8,c('Monocular','Binocular','Dichoptic','Dichoptic cross'),lwd=3,col=colpal[condlist],box.lwd=2)

text(17,0.95,'(g)',adj=0.5,cex=2.5)

@@

legend(62,95,c('48% match','24% match'),col=colpal[2:1],lwd=3,pt.cex=2,box.lwd=2)

@@

text(10,117.5,'(h)',adj=0.5,cex=2.5)

postscript("pdata.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

for (cond in 1:3){
  lines(contrastsdB,abs(cleanmeansP[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  points(contrastsdB,abs(cleanmeansP[cond,]), pch = 20+cond, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves
}  
cond <- 6
  lines(contrastsdB,abs(cleanmeansP[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  points(contrastsdB,abs(cleanmeansP[cond,]), pch = 25, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves

plotlims <- c(15,40,0,1)  
ticklocsx <- contrastsdB    # locations of tick marks on x axis
ticklocsy <- seq(0,1,0.2)    # locations of tick marks on y axis
ticklabelsx <-c(6,12,24,48,96)        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

postscript("edata.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Amplitude (µV) at 2Hz", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

for (cond in 1:3){
  lines(contrastsdB,abs(cleanmeansEav[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  points(contrastsdB,abs(cleanmeansEav[cond,]), pch = 20+cond, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves
}  
cond <- 6
  lines(contrastsdB,abs(cleanmeansEav[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  points(contrastsdB,abs(cleanmeansEav[cond,]), pch = 25, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves

# legend(22,0.8,c('Monocular','Binocular','Dichoptic','Dichoptic cross'),pch=c(21:23,25),pt.bg=colpal[c(1:3,6)],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(17,0.95,'(b)',adj=0.5,cex=2.5)

postscript("edata2.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

par(pty="s")  # make axis square
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1) 
title(xlab="Target contrast (%)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)  
title(ylab="Amplitude (µV) at 4Hz", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

for (cond in 1:3){
  lines(contrastsdB,abs(cleanmeansE2av[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  points(contrastsdB,abs(cleanmeansE2av[cond,]), pch = 20+cond, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves
}  
cond <- 6
  lines(contrastsdB,abs(cleanmeansE2av[cond,]), col=colpal[cond], lwd=3, cex=0.5)
  points(contrastsdB,abs(cleanmeansE2av[cond,]), pch = 25, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves

# legend(22,0.8,c('Monocular','Binocular','Dichoptic','Dichoptic cross'),pch=c(21:23,25),pt.bg=colpal[c(1:3,6)],pt.lwd=3,pt.cex=1.6,box.lwd=2)
text(17,0.95,'(c)',adj=0.5,cex=2.5)

postscript("matchingdata.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5.5, width = 5.5)

plotlims <- c(0,125,0,125)  
ticklocs <- seq(0,125,25)   
ratiosL <- c(1, 1, 1, 1, 1, 0.75, 0.5, 0.25, 0)
ratiosR <- c(0, 0.25, 0.5, 0.75, 1, 1, 1, 1, 1)

  # if ((SEthresh[matchlev,cond]+meanthresh[matchlev,cond])<42){
  # arrows(ratiosR[cond]*10^(meanthresh[matchlev,cond]/20),ratiosL[cond]*10^(meanthresh[matchlev,cond]/20),ratiosR[cond]*10^((SEthresh[matchlev,cond]+meanthresh[matchlev,cond])/20),ratiosL[cond]*10^((SEthresh[matchlev,cond]+meanthresh[matchlev,cond])/20),length=0.015, angle=90, lwd=2)
  # }
  # 
  # # plot as an arrow if exceeding the axis limits
  # if ((SEthresh[matchlev,cond]+meanthresh[matchlev,cond])>42){
  # arrows(ratiosR[cond]*10^(meanthresh[matchlev,cond]/20),ratiosL[cond]*10^(meanthresh[matchlev,cond]/20),125,ratiosL[cond]*10^((SEthresh[matchlev,cond]+meanthresh[matchlev,cond])/20),length=0.035, angle=45, lwd=2)
  # }

  # arrows(ratiosR[cond]*10^(meanthresh[matchlev,cond]/20),ratiosL[cond]*10^(meanthresh[matchlev,cond]/20),ratiosR[cond]*10^((meanthresh[matchlev,cond]-SEthresh[matchlev,cond])/20),ratiosL[cond]*10^((meanthresh[matchlev,cond]-SEthresh[matchlev,cond])/20),length=0.015, angle=90, lwd=2)
}
}

@@

  PostScriptTrace('pdata.ps')
  p6 <- readPicture('pdata.ps.xml')
  PostScriptTrace('edata.ps')
  p7 <- readPicture('edata.ps.xml')
  PostScriptTrace('edata2.ps')
  p8 <- readPicture('edata2.ps.xml')
  PostScriptTrace('matchingdata.ps')
  p9 <- readPicture('matchingdata.ps.xml')

@@

  pdf(paste0(figdir,"modelfigure.pdf"), bg="transparent", height = 10, width = 30)
  par(mar=c(0.1,0.1,0.1,0.1))
  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,1))

  grid.picture(p1,x=1/12,y=0.25,width=1/6,height=1)
  grid.picture(p2,x=3/12,y=0.25,width=1/6,height=1)
  grid.picture(p3,x=5/12,y=0.25,width=1/6,height=1)
  grid.picture(p4,x=7/12,y=0.25,width=1/6,height=1)

  grid.picture(p5,x=5/6,y=0.5,width=1/3,height=1)

  grid.picture(p6,x=1/12,y=0.75,width=1/6,height=1)
  grid.picture(p7,x=3/12,y=0.75,width=1/6,height=1)
  grid.picture(p8,x=5/12,y=0.75,width=1/6,height=1)
  grid.picture(p9,x=7/12,y=0.75,width=1/6,height=1)

@@

  file.remove(c('CRF1e.ps','CRF2e.ps','CRF1p.ps','matchingmodel.ps','matchingdata.ps','pdata.ps','edata.ps','edata2.ps','posteriors.ps'))
  file.remove(c('CRF1e.ps.xml','CRF2e.ps.xml','CRF1p.ps.xml','matchingmodel.ps.xml','matchingdata.ps.xml','pdata.ps.xml','edata.ps.xml','edata2.ps.xml','posteriors.ps.xml'))

>>>>

text(9,0.58,'(e)',cex=2.5,adj=0.5)

@@

for (cond in 1:3){
  points(contrastsdB,abs(cleanmeansP[cond,]), pch = 20+cond, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves
}  
cond <- 6
  points(contrastsdB,abs(cleanmeansP[cond,]), pch = 25, col='black', bg=colpal[cond], cex=1.6, lwd=3)

@@

for (cond in 1:3){
  points(contrastsdB,abs(cleanmeansEav[cond,]), pch = 20+cond, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves
}  
cond <- 6
  points(contrastsdB,abs(cleanmeansEav[cond,]), pch = 25, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points

text(17,0.95,'(b)',adj=0.5,cex=2.5)

@@

for (cond in 1:3){
  points(contrastsdB,abs(cleanmeansE2av[cond,]), pch = 20+cond, col='black', bg=colpal[cond], cex=1.6, lwd=3)   # draw the data points themselves
}  
cond <- 6
  points(contrastsdB,abs(cleanmeansE2av[cond,]), pch = 25, col='black', bg=colpal[cond], cex=1.6, lwd=3)

text(17,0.95,'(c)',adj=0.5,cex=2.5)

@@

for (matchlev in 1:2){
for (cond in 1:9){
  points(ratiosR[cond]*10^(meanthresh[matchlev,cond]/20),ratiosL[cond]*10^(meanthresh[matchlev,cond]/20),pch=16,cex=2,col=colpal[matchlev])
}}

legend(62,95,c('48% match','24% match'),col=colpal[2:1],lwd=3,pch=16,pt.cex=2,box.lwd=2)

@@

@@

@@

  pdf(paste0(figdir,"modelfigure.pdf"), bg="transparent", height = 5, width = 25)
  par(mar=c(0.1,0.1,0.1,0.1))
  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,1))

  grid.picture(p1,x=1/10,y=0.5,width=0.2,height=1)
  grid.picture(p2,x=3/10,y=0.5,width=0.2,height=1)
  grid.picture(p3,x=5/10,y=0.5,width=0.2,height=1)
  grid.picture(p4,x=7/10,y=0.5,width=0.2,height=1)
  grid.picture(p5,x=9/10,y=0.5,width=0.2,height=1)

@@

  file.remove(c('CRF1e.ps','CRF2e.ps','CRF1p.ps','matchingmodel.ps','posteriors.ps'))
  file.remove(c('CRF1e.ps.xml','CRF2e.ps.xml','CRF1p.ps.xml','matchingmodel.ps.xml','posteriors.ps.xml'))

### 2023-06-26 manuscript.tex
Simplified presentation of model figure

In the steady-state literature, one hallmark of a nonlinear system that pools inputs is the presence of intermodulation responses at the sums and differences of the fundamental flicker frequencies (\protect\hyperlink{ref-Tsai2012}{Tsai et al., 2012}). In Figure \ref{fig:IMfig} we plot the amplitude spectra of conditions from Experiment 1 in which the two eyes were stimulated at different frequencies (2Hz and 1.6Hz) but at the same contrast (48\%; these correspond to the binocular cross and dichoptic cross conditions in Figures \ref{fig:pupildata}d,e and \ref{fig:EEGdata}d,e). Figure \ref{fig:IMfig}a reveals a strong intermodulation difference response at 0.4Hz (red dashed line), and Figure \ref{fig:IMfig}b reveals an intermodulation sum response at 3.6Hz (red dashed line). It seems likely that the absence of a sum response for pupillometry data, and of a difference responses for the EEG data, is a consequence of the temporal constraints of these methods. The presence of intermodulation terms is predicted by nonlinear gain control models of the type considered here (\protect\hyperlink{ref-Baker2017}{Baker and Wade, 2017}; \protect\hyperlink{ref-Tsai2012}{Tsai et al., 2012}), and indicates that the processing of monocular flicker signals is not fully linear prior to the point at which they are combined across the eyes. Indeed, our model architecture (\protect\hyperlink{ref-Meese2006}{Meese et al., 2006}) makes specific predictions about the location of interocular suppression - it impacts before binocular combination, consistent with results from primate physiology (\protect\hyperlink{ref-Dougherty2019}{Dougherty et al., 2019}).

@@

>>>>

In the steady-state literature, one hallmark of a nonlinear system that pools inputs is the presence of intermodulation responses at the sums and differences of the fundamental flicker frequencies (\protect\hyperlink{ref-Baitch1988}{Baitch and Levi, 1988}; \protect\hyperlink{ref-Tsai2012}{Tsai et al., 2012}). In Figure \ref{fig:IMfig} we plot the amplitude spectra of conditions from Experiment 1 in which the two eyes were stimulated at different frequencies (2Hz and 1.6Hz) but at the same contrast (48\%; these correspond to the binocular cross and dichoptic cross conditions in Figures \ref{fig:pupildata}d,e and \ref{fig:EEGdata}d,e). Figure \ref{fig:IMfig}a reveals a strong intermodulation difference response at 0.4Hz (red dashed line), and Figure \ref{fig:IMfig}b reveals an intermodulation sum response at 3.6Hz (red dashed line). It seems likely that the absence of a sum response for pupillometry data, and of a difference responses for the EEG data, is a consequence of the temporal constraints of these methods. The presence of intermodulation terms is predicted by nonlinear gain control models of the type considered here (\protect\hyperlink{ref-Baker2017}{Baker and Wade, 2017}; \protect\hyperlink{ref-Tsai2012}{Tsai et al., 2012}), and indicates that the processing of monocular flicker signals is not fully linear prior to the point at which they are combined across the eyes. Indeed, our model architecture (\protect\hyperlink{ref-Meese2006}{Meese et al., 2006}) makes specific predictions about the location of interocular suppression - it impacts before binocular combination, consistent with results from primate physiology (\protect\hyperlink{ref-Dougherty2019}{Dougherty et al., 2019}).

@@

\leavevmode\vadjust pre{\hypertarget{ref-Baitch1988}{}}%
Baitch LW, Levi DM. 1988. Evidence for nonlinear binocular interactions in human visual cortex. \emph{Vision Res} \textbf{28}:1139--43. doi:\href{https://doi.org/10.1016/0042-6989(88)90140-x}{10.1016/0042-6989(88)90140-x}
